Mamary Gland Development Exploratory Data Analysis

### 2017-01-12 src/QCAnalysis.Rmd
Updated QC_Analysis, less conservative and included DimRed plots

---
title: "Mamary Gland Development Exploratory Data Analysis"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
---
***
# Load Data and Packages
```{r, message=FALSE}
library(scran)
library(dplyr)
library(knitr)
library(ggplot2)
source("functions.R")
dataList <- readRDS("../data/Robjects/ExpressionList.rds")
m <- dataList[["counts"]]
pD <- dataList[["phenoData"]]
fD <- dataList[["featureData"]]

```
# Overview over the data

@@

## QC Thresholding
```{r}
#simple thresholds
MitoCutOff <- 0.05
genesDetectedCutOff <- 1000
LibSizeCutOff <- 2000
```
Four different thresholds were applied:

@@

kable(t(as.matrix(table(pD$Condition,pD$PassAll))),caption="Cells passing QC in different Conditions")
#Apply Thresholds
pD <- filter(pD, PassAll)
m <- m[,as.character(pD$barcode)]

#Thresholding on lowly expressed genes
expThreshold <- 0.1
keep <- rowMeans(m) > expThreshold
m <- m[keep,]
fD <- fD[keep,]
#Final matrix
dim(m)
```

## Normalization and first visualization
```{r}
clusters <- quickCluster(m)
pD$sf <- computeSumFactors(m,clusters=clusters,positive=TRUE)
m.norm <- t(t(m)/pD$sf)
# m.norm <- m.norm[,!is.na(colSums(m.norm))]
pD <- filter(pD, barcode %in% colnames(m.norm))
# 
fD <- mutate(fD, meanExpression=rowMeans(m.norm),
	     CV2=apply(m.norm,1,cv2),
	     dm=DM(meanExpression,CV2))

highVar <- dplyr::arrange(fD,desc(dm)) %>% .$id
fD$highVar <- fD$id %in% highVar[1:1000]
VarPlot <- ggplot(fD, aes(x=log10(meanExpression),y=log10(CV2), color=highVar)) +
    geom_point() +
    theme_bw()

fPCA <- m.norm[fD$highVar,]
pc <- prcomp(log2(t(fPCA)+1))
tsn <- Rtsne(log2(t(fPCA)+1),max_iter=5000)

pD$PC1 <- pc$x[,1]
pD$PC2 <- pc$x[,2]
pD$tSNE1 <- tsn$Y[,1]
pD$tSNE2 <- tsn$Y[,2]

Acta2 <- filter(fD, symbol=="Acta2") %>% .$id
pD$Acta2 <- t(m.norm)[,Acta2]
# 
pcaPlot <- ggplot(pD, aes(x=PC1, y=PC2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=4) +
    theme_bw() 
pcaPlot

tsnePlot <- ggplot(pD, aes(x=tSNE1, y=tSNE2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=4) +
    theme_bw()
tsnePlot
```

>>>>

---
title: "Mamary Gland Development Exploratory Data Analysis"
author: "Karsten Bach"
date: '`r Sys.Date()`'
output:
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: united
    highlight: tango
---
***
# Load Data and Packages
```{r, message=FALSE}
library(scran)
library(dplyr)
library(knitr)
library(ggplot2)
rnd_seed <- 300
library(Rtsne)
source("functions.R")
dataList <- readRDS("../data/Robjects/ExpressionList.rds")
m <- dataList[["counts"]]
pD <- dataList[["phenoData"]]
fD <- dataList[["featureData"]]
```
# Overview over the data

@@

## QC Thresholding
```{r}
#simple thresholds
MitoCutOff <- 0.05
genesDetectedCutOff <- 500
LibSizeCutOff <- 500
```
Four different thresholds were applied:

@@

kable(t(as.matrix(table(pD$Condition,pD$PassAll))),caption="Cells passing QC in different Conditions")
#Apply Thresholds
pD.filtered <- filter(pD, PassAll)
m.filtered <- m[,as.character(pD.filtered$barcode)]

#Thresholding on lowly expressed genes
expThreshold <- 0.1
keep <- rowMeans(m.filtered) > expThreshold
m.filtered <- m.filtered[keep,]
fD.filtered <- fD[keep,]
#Final matrix
dim(m.filtered)
```

## Normalization and first visualization
```{r}
clusters <- quickCluster(m.filtered)
pD.filtered$sf <- computeSumFactors(m.filtered,clusters=clusters,positive=TRUE)
m.norm <- t(t(m.filtered)/pD.filtered$sf)
m.norm <- m.norm[,!is.na(colSums(m.norm))]
pD.filtered <- filter(pD.filtered, barcode %in% colnames(m.norm))
# 
fD.filtered <- mutate(fD.filtered, meanExpression=rowMeans(m.norm),
	     CV2=apply(m.norm,1,cv2),
	     dm=DM(meanExpression,CV2))

highVar <- dplyr::arrange(fD.filtered,desc(dm)) %>% .$id
topX <- 0.1*length(highVar)
fD.filtered$highVar <- fD.filtered$id %in% highVar[1:topX]
VarPlot <- ggplot(fD.filtered, aes(x=log10(meanExpression),y=log10(CV2), color=highVar)) +
    geom_point() +
    theme_bw()
VarPlot

fPCA <- log2(t(m.norm[fD.filtered$highVar,])+1)
fPCA <- scale(fPCA,scale=TRUE,center=TRUE)
pc <- prcomp(fPCA)
set.seed(rnd_seed)
tsn <- Rtsne(fPCA,perplexity=floor(nrow(fPCA)/5))

pD.filtered$PC1 <- pc$x[,1]
pD.filtered$PC2 <- pc$x[,2]
pD.filtered$tSNE1 <- tsn$Y[,1]
pD.filtered$tSNE2 <- tsn$Y[,2]

Acta2 <- filter(fD.filtered, symbol=="Acta2") %>% .$id
pD.filtered$Acta2 <- t(m.norm)[,Acta2]
# 
pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

pcaPlot <- ggplot(pD.filtered, aes(x=PC1, y=PC2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw() 
pcaPlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=Condition, shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot

tsnePlot <- ggplot(pD.filtered, aes(x=tSNE1, y=tSNE2, color=log2(Acta2+1), shape=Replicate)) +
    geom_point(alpha=0.8,size=2) +
    theme_bw()
tsnePlot
```
## Save Data
```{r}
out <- list()
out[[1]] <- m
out[[2]] <- pD
out[[3]] <- fD
saveRDS(out,file="../data/Robjects/ExpressionList_QC.rds")
```
